Multilevel generalizations of collective spin dynamics and squeezing

### 2019-11-05 collective_sun/multilevel_spin_notes.tex
added section on a periodic drive

\newcommand{\B}{\mathcal{B}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\E}{\mathcal{E}}
\renewcommand{\H}{\mathcal{H}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\M}{\mathcal{M}}
\newcommand{\N}{\mathcal{N}}
\renewcommand{\O}{\mathcal{O}}
\renewcommand{\P}{\mathcal{P}}
\newcommand{\Q}{\mathcal{Q}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\T}{\mathcal{T}}
\renewcommand{\S}{\mathcal{S}}
\newcommand{\V}{\mathcal{V}}
\newcommand{\X}{\mathcal{X}}
\newcommand{\Z}{\mathcal{Z}}

@@

In addition to the fact that the fully symmetric manifold $\M_\FS$ is
ground-state manifold of the interaction Hamltonian $H_{\t{int}}$, an
important feature of $\M_\FS$ is its closure under the action of
collective operators of the form $\Q=\sum_pQ^{(p)}$, as well as linear
combinations and products thereof.  Closure of $\M_\FS$ under
collective the action of collective operators, and in particular
collective dynamics, is a straightforward consequence of the
permutational symmetry obeyed by collective operators.

@@

@@

\begin{itemize}
\item Visualization of multilevel states
  \begin{itemize}
  \item Single-particle eigenstates of drive operators
  \item Single-particle dynamics induced by drive operators
  \item Collective states and dynamics
  \end{itemize}
\item Classifying eigenstates of SU($n$)-symmetric interactions
\item Implications of SOC
  \begin{itemize}
  \item Special angles $\phi$
  \end{itemize}
\item Generalizations of squeezing beyond SU(2)?
\item Consider different interactions
  \begin{itemize}
  \item Include electronic states
  \item Super-exchange regime
  \item Non-uniform (``disordered'') interactions
  \end{itemize}
\item Consider different drives
\end{itemize}

>>>>

\newcommand{\B}{\mathcal{B}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\E}{\mathcal{E}}
\renewcommand{\H}{\mathcal{H}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\J}{\mathcal{J}}
\newcommand{\M}{\mathcal{M}}
\newcommand{\N}{\mathcal{N}}
\renewcommand{\O}{\mathcal{O}}
\renewcommand{\P}{\mathcal{P}}
\newcommand{\Q}{\mathcal{Q}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\T}{\mathcal{T}}
\renewcommand{\S}{\mathcal{S}}
\newcommand{\V}{\mathcal{V}}
\newcommand{\X}{\mathcal{X}}
\newcommand{\Z}{\mathcal{Z}}

@@

In addition to the fact that the fully symmetric manifold $\M_\FS$ is
ground-state manifold of the interaction Hamltonian $H_{\t{int}}$, an
important feature of $\M_\FS$ is its closure under the action of
collective operators of the form $\Q=\sum_pQ^{(p)}$, as well as linear
combinations and products thereof.  Closure of $\M_\FS$ under
collective the action of collective operators, and in particular under
collective dynamics, is a straightforward consequence of the
permutational symmetry obeyed by these operators.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Periodic drive}

We now consider the effect of applying periodic drive of the form
\begin{align}
  H_{\t{pd}}\p{t} = \sum_p \chi_p \omega \cos\p{\omega t} D_\z^{(p)},
\end{align}
where $\omega$ is a driving frequency, and $\chi_p$ is a (real)
dimensionless driving amplitude for spin $p$, and
$D_\z^{(p)} \equiv \xi_1 D_{1,0}^{(p)}$ with $\xi_1$ defined in
\eqref{eq:scale_fac}.  The periodic drive $H_{\t{pd}}\p{t}$ can be
implemented directly using external magnetic fields or driving lasers.
Transition operators $T_{LM}$ transform under rotations generated by
the drive operator $D_\z$ as
\begin{align}
  e^{i\phi D_\z} T_{LM} e^{-i\phi D_\z} = e^{iM\phi} T_{LM}.
\end{align}
If we move into the rotating frame of the drive, then the interaction
Hamiltonian transforms as
\begin{align}
  H_{\t{int}} \to \tilde H_{\t{int}}
  \equiv U_{\t{pd}}\p{t}^\dag H_{\t{int}} U_{\t{pd}}\p{t}
\end{align}
where
\begin{align}
  U_{\t{pd}}\p{t} \equiv \exp\sp{-i\int_0^td\tau\,H_{\t{pd}}\p{\tau}}
  = \prod_p \exp\sp{-i \chi_p \sin\p{\omega t} D_\z^{(p)}}.
\end{align}
Expanding the interaction Hamiltonian $H_{\t{int}}$ in terms of
transition operators $T_{LM}$, we find that
\begin{align}
  \tilde H_{\t{int}}
  = -\f{u}{2} \sum_{p,q,L,M} e^{iM\p{\chi_q-\chi_p}\sin\p{\omega t}}
  T_{LM}^{(p)\dag} T_{LM}^{(q)}
  = -\f{u}{2} \sum_{p,q,L,M,n}
  \J_n\p{M\sp{\chi_q-\chi_p}} e^{in\omega t}
  T_{LM}^{(p)\dag} T_{LM}^{(q)},
\end{align}
where $\J_n$ is the $n$-th order Bessel function of the first kind.
If the drive frequency $\omega$ is much larger than the interaction
strength $u$, i.e.~$\omega\gg u$, then we can neglect oscillating
terms with $n\ne0$, arriving at
\begin{align}
  \tilde H_{\t{int}} \approx -\f{u}{2} \sum_{p,q,L,M}
  \J_0\p{M\sp{\chi_q-\chi_p}}
  T_{LM}^{(p)\dag} T_{LM}^{(q)}.
\end{align}

\vspace{5cm}

@@

\begin{itemize}
\item Visualization of multilevel states
  \begin{itemize}
  \item Single-particle eigenstates of drive operators
  \item Single-particle dynamics induced by drive operators
  \item Collective states and dynamics
  \end{itemize}
\item Implications of SOC
  \begin{itemize}
  \item Special angles $\phi$
  \end{itemize}
\item Generalizations of squeezing beyond SU(2)?
\item Consider different interactions
  \begin{itemize}
  \item Include electronic states
  \item Super-exchange regime
  \item Non-uniform (``disordered'') interactions
  \end{itemize}
\item Consider different drives
\end{itemize}
